Supplement B: Stationwise Evaluation

### 2019-02-07 pkg/disttree/inst/draft/draft.Rnw
first revisions based on reviews; documentation of changes in rejoinder2.Rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2448 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: b45f4b30e679e49be7210ebdd93fc64574908019

\item \emph{Distributional forest:} All 80~predictor variables are
  considered for learning a forest of 100 trees. Subsampling is employed 
  for each tree using a third of the predictors
  in each split of the tree (\code{mtry}). Parameters are estimated by
  adaptive local likelihood based on the forest weights as described
  in Section~\ref{sec:methods}. 
  The stopping criteria are the minimal number of observations to perform a split 
  (\texttt{minsplit} $= 50$), the minimal number of observations in a 
  segment (\texttt{minbucket} $= 20$), and the significance level 
  for variable selection (\texttt{alpha} $= 1$). The latter means
  that no early stopping (or ``pre-pruning'') is applied
  based on results of the statistical tests.
  % stopping criteria:
  % forest_minsplit = 50, forest_minbucket = 20, forest_mincrit = 0

@@

In a first step, we show a detailed comparison of the competing models
for one observation site, Axams in Tyrol (in the center of the study area,
see Figure~\ref{fig:all-map}). As for all other stations, daily precipitation 
observations and numerical weather predictions are available for the month of July 
from 1985 through 2012. In Figure~\ref{fig:axams-dist} in the introduction the 
probabilistic forecasts from the distributional forest, trained on 1985--2008, 
for July 24 in 2009--2012 have already been shown as a motivational example. 
The figure shows that the model properly depicts the point masses at zero 
(i.e., the probability of a dry day) and the forecasted probability density 
function for the total amount of precipitation. The four sample forecasts differ 
considerably in location $\mu$, scale $\sigma$, and the amount of
censoring while conforming quite well with the actual observations
from these days. While this is a nice illustrative example we are  
interested in the overall predictive performance and calibration of the
distributional fits. More details of this assessment as well as
an application to 14~further meteorological stations is provided in
Supplement~B \citep{Schlosser+Hothorn+Stauffer:2018b}.

@@

\subsubsection{Distributional specifications}

@@

\section*{Computational details}

@@

@@

With the conditional expectation $\mu_l$ and the covariance 
$\Sigma_l$ of $T_l$ as derived by 
\cite{Strasser+Weber:1999} the test statistic can be standardized. 
The observed multivariate linear statistic $t_l$ which is either an
$k$~-~or~$k \cdot H$~-~dimensional vector, depending on the scale of $Z_l$,
is mapped onto the real line by a univariate test statistic $c$.
In the application of this paper a quadratic form is chosen, such that
\begin{equation}
c_{\text{quad}}(t_l,\mu_l,\Sigma_l) = (t_l-\mu_l)\Sigma_l^+(t_l-\mu_l)^{\top}
\end{equation}
where $\Sigma_l^+$ is the Moore-Penrose inverse of $\Sigma_l$. Alternatively,
the maximum of the absolute values of the standardized linear statistic
can be considered.

>>>>

\item \emph{Distributional forest:} All 80~predictor variables are
  considered for learning a forest of 100 trees. Subsampling is employed 
  for each tree using a third of the predictors
  in each split of the tree (\hlc{specified by the argument} \code{mtry} 
  \hlc{in the applied function} \code{distforest}; 
  \hlc{more information on the employed software is provided in Section}
  ``Computational Details''). 
  Parameters are estimated by
  adaptive local likelihood based on the forest weights as described
  in Section~\ref{sec:methods}. 
  The stopping criteria are the minimal number of observations to perform a split 
  (\texttt{minsplit} $= 50$), the minimal number of observations in a 
  segment (\texttt{minbucket} $= 20$), and the significance level 
  for variable selection (\texttt{alpha} $= 1$). The latter means
  that no early stopping (or ``pre-pruning'') is applied
  based on results of the statistical tests.
  % stopping criteria:
  % forest_minsplit = 50, forest_minbucket = 20, forest_mincrit = 0

@@

In a first step, we show a detailed comparison of the competing models
for one observation site, Axams in Tyrol (in the center of the study area,
see Figure~\ref{fig:all-map}). As for all other stations, daily precipitation 
observations and numerical weather predictions are available for the month of July 
from 1985 through 2012. In Figure~\ref{fig:axams-dist} in the introduction the 
probabilistic forecasts from the distributional forest, trained on 1985--2008, 
for July 24 in 2009--2012 have already been shown as a motivational example. 
\hlc{In particular, the figure displays the point mass at zero (i.e., the probability 
of a dry day) depicted by the model together with the forecasted probability 
density function for the total amount of precipitation. Based on this illustration 
it can be observed that} the four sample forecasts differ considerably in 
location $\mu$, scale $\sigma$, and the amount of censoring while conforming quite 
well with the actual observations from these days. 
While this is a nice illustrative example we are  
interested in the overall predictive performance and calibration of the
distributional fits. More details of this assessment as well as
an application to 14~further meteorological stations is provided in
Supplement~B \citep{Schlosser+Hothorn+Stauffer:2018b}.

@@

\subsubsection*{Distributional specifications}

@@

\section*{Computational details}
\label{sec:comp}

@@

\section*{Acknowledgements}
\hlc{An extended research stay of Torsten Hothorn in Innsbruck (August
2017 to January 2018) was financially supported by the Swiss National
Science Foundation, grant number SNF IZSEZ0 177091.}

@@

With the conditional expectation $\mu_l$ and the covariance 
$\Sigma_l$ of $T_l$ as derived by 
\cite{Strasser+Weber:1999} the test statistic can be standardized. 
The observed multivariate linear statistic $t_l$ which is either an
$k$~-~or~$k \cdot H$~-~dimensional vector, depending on the scale of $Z_l$,
is mapped onto the real line by a univariate test statistic $c$.
In the application of this paper a quadratic form is chosen, such that
\begin{equation}
c_{\text{quad}}(t_l,\mu_l,\Sigma_l) = (t_l-\mu_l)\Sigma_l^+(t_l-\mu_l)^{\top}
\end{equation}
where $\Sigma_l^+$ is the Moore-Penrose inverse of $\Sigma_l$. Alternatively,
the maximum of the absolute values of the standardized linear statistic
can be considered ($c_{\text{max}}$).

\hlc{Since} \cite{Strasser+Weber:1999} \hlc{showed that the asymptotic conditional 
distribution of a linear statistics with conditional expectation} $\mu$ 
\hlc{and covariance} $\Sigma$ 
\hlc{is a multivariate normal distribution with parameters}
$\mu$ and $\Sigma$ 
\hlc{the asymptotic conditional distribution of the standardized test statistic}
$c(t_{l},\mu_{l},\Sigma_{l})$ 
\hlc{is either normal}
($c_{\text{max}}$) 
\hlc{or follows a} 
$\chi^2$ distribution ($c_{\text{quad}}$).

### 2019-02-07 pkg/disttree/inst/draft/supplement/supplement_stationwise.Rnw
final formatting
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2378 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 4503d2d73b067eabaa07a9184943d75e94a802fc
first revisions based on reviews; documentation of changes in rejoinder2.Rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2448 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: b45f4b30e679e49be7210ebdd93fc64574908019

\usepackage{graphicx, natbib}
\pagestyle{empty}
\usepackage{Sweave}
\usepackage{amstext,amsfonts,amsmath,bm,thumbpdf,lmodern,pdfpages,hyperref,bookmark}

@@

\Abstract{
  In the case study presented in the main manuscript ``Distributional Regression
  Forests for Probabilistic Precipitation Forecasting in Complex Terrain''
  the novel distributional regression forests are evaluated and compared with
  other distributional regression approaches on 95 meteorological stations.
  Moreover, further details -- including a cross-validation analysis -- are
  preformed for only one station, Axams. To show that Axams is a fairly typical
  station and similar insights can be obtained for other stations as well,
  this supplement presents the same analysis as in Section~3.3 of the main
  manuscript for 14~further meteorological stations.
}

@@

The Resiudal QQ plots and the PIT histograms visualize the same values,
however, on different scales. For each observation $y$ of the test sample 
the corresponding PIT value is calculated by applying the probability function 
$F(y; \hat{\mu}, \hat{\sigma})$
with the estimated distribution parameters plugged in. The resulting values
can either be plotted in a histogram (PIT histogram) or be plotted against
the corresponding quantiles of a standard normal distribution (residual QQ plot)
for a continuous variable $Y$. To account for the point mass at 0, for each 
observation equal to 0 the corresponding PIT value is multiplied with a randomized
value following a uniform distribution on $[0,1]$. In that way these values
are apportioned on those bins of the histogram that would not contain any
values because no observation is beneath 0.

>>>>

\usepackage{graphicx, natbib}
\pagestyle{empty}
\usepackage{Sweave}
\usepackage{amstext,amsfonts,amsmath,bm,thumbpdf,lmodern,pdfpages,hyperref,bookmark}
\usepackage[all]{hypcap}

@@

\Abstract{
  In the case study presented in the main manuscript ``Distributional Regression
  Forests for Probabilistic Precipitation Forecasting in Complex Terrain''
  the novel distributional regression forests are evaluated and compared with
  other distributional regression approaches on 95 meteorological stations.
  Moreover, further details -- including a cross-validation analysis -- are
  performed for only one station, Axams. To show that Axams is a fairly typical
  station and similar insights can be obtained for other stations as well,
  this supplement presents the same analysis as in Section~3.3 of the main
  manuscript for 14~further meteorological stations.
}

@@

The Residual QQ plots and the PIT histograms visualize the same values,
however, on different scales. For each observation $y$ of the test sample 
the corresponding PIT value is calculated by applying the probability function 
$F(y; \hat{\mu}, \hat{\sigma})$
with the estimated distribution parameters plugged in. The resulting values
can either be plotted in a histogram (PIT histogram) or be plotted against
the corresponding quantiles of a standard normal distribution (residual QQ plot)
for a continuous variable $Y$. To account for the point mass at 0, for each 
observation equal to 0 the corresponding PIT value is multiplied with a randomized
value following a uniform distribution on $[0,1]$. In that way these values
are apportioned on those bins of the histogram that would not contain any
values because no observation is beneath 0.
